Proof-Reconstruction in Type Theory for Propositional Logic

### 2017-10-23 paper/sections/proof-reconstruction/strip.tex
[ paper ] last changes.

\begin{mainlemma}
  \label{lem:inv-uh-lem}
  If $Γ ⊢ \fuh\ φ$ then $Γ ⊢ φ$.
\end{mainlemma}

\begin{sketchproof}
Use induction on the cases defined by the outcome of the
$\fuh$ function in~(\ref{eq:uh-structured}).
\begin{itemize}
  \item If $n = 0$, by definition we conclude $Γ ⊢ φ$.
  \item If $n = 1$, we apply the following theorem proved in~\cite{AgdaProp},
    \begin{equation*}
    \texttt{∧⇒\rm{-to-}⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
    \end{equation*}
  \item For $n > 1$, we use induction on the structure of the second
        argument.
\vskip 2mm
\begin{itemize}

@@

Finally, using the theorem \texttt{⇒∧⇒\rm{-to-}⇒∧} from~\cite{AgdaProp},
\begin{equation*}
  \texttt{⇒∧⇒\rm{-to-}⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃),
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}_1$}
  \AxiomC{$\mathcal{D}_2$}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
  \RightLabel{\tt ⇒∧⇒\rm{-to-}⇒∧}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)}
  \end{bprooftree}
\end{equation*}
\item Other cases are proved in a similar way.
\end{itemize}
\end{itemize}
\end{sketchproof}

@@

%  1137  git add .
%  1138  git commit -am "[ paper ] strip proof."
%  1139  git pull origin paper
%  1140  git push origin paper
\begin{equation}
\label{eq:strip-fixed}
\begin{aligned}
&\hspace{.495mm}\fstrip : \Prop → \Nat → \Prop\\
&\begin{array}{lll}
\fstrip &(φ₁ ∧ φ₂)     ~(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ∨ φ₂)     ~(\suc~n) &= \fuh~(¬ φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ⇒ φ₂)     ~(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(¬ (φ₁ ∧ φ₂)) ~(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ∨ φ₂)) ~(\suc~n) &= \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ⇒ φ₂)) ~(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ ¬ φ₁)      ~(\suc~n) &= \fuh~(\fstrip~φ₁~n)\\
\fstrip &(¬ ⊥)         ~(\suc~n) &= ⊤\\
\fstrip &(¬ ⊤)         ~(\suc~n) &= ⊥\\
\fstrip &φ             ~\text{n} &= φ
\end{array}
\end{aligned}
\end{equation}

\begin{mainlemma}
\label{lem:lem-inv-strip}
Let $\text{n} : \Nat$ the complexity measure of the strip function
in~(\ref{eq:strip-unfixed}).
If $Γ ⊢ \fstrip~φ~n$ then $Γ ⊢ φ$.
\end{mainlemma}

\begin{sketchproof}
The proof is by induction on the structure of the
formula~$φ$ by following the cases in~(\ref{eq:strip-fixed}).
We present a straightforward case with double negation, a case with
conjunction connective, and last, the case with a negated disjunction.
We refer the reader to~\cite{AgdaMetis} for the complete proof in \Agda.

Case $φ ≡ ¬ ¬ φ₁$.
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \fstrip~(¬¬φ₁)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n)$}
  \RightLabel{Lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ \fstrip~φ₁~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

Case $φ ≡ φ₁ ∧ φ₂$. We get a proof for each conjunct and using the
introduction rule for conjunction connective, the expected result follows.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fstrip~(φ₁ ∧ φ₂)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n)$}
  \RightLabel{Lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ \fstrip~φ₁~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ φ₁$}
    \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \fstrip~(φ₁ ∧ φ₂)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{Lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip~φ₂~n$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \fstrip~φ₂~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ φ₂$}
  \end{bprooftree}
\end{equation*}

Case $φ ≡ ¬ (φ₁ ∨ φ₂)$. We show the theorems $Γ ⊢ ¬ φ₁$ and $Γ ⊢ ¬ φ₂$.
With the conjunction of $¬ φ₁$ and $¬ φ₂$ by applying De Morgan Law, the result follows.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \fstrip~(¬ (φ₁ ∨ φ₂))~(\suc~n)$}
\RightLabel{by def.}
\UnaryInfC{$Γ ⊢ \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ \fuh~(\fstrip~¬ φ₁~n)$}
\RightLabel{Lemma~\ref{lem:inv-uh-lem}}
\UnaryInfC{$Γ ⊢ \fstrip~¬ φ₁~n$}
\RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ ¬φ₁$}
\end{bprooftree}
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{$Γ ⊢ ¬ φ₁$}
  \AxiomC{$Γ ⊢ \fstrip~(¬ (φ₁ ∨ φ₂))~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)$}
  \RightLabel{∧-proj₂}
  \UnaryInfC{$Γ ⊢ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)$}
  \RightLabel{Lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \fstrip~¬ φ₂~n$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ ⊢ \fstrip~¬ φ₂~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ ¬ φ₂$}
  \end{bprooftree}
\end{equation*}
\end{sketchproof} % the proof continue...

@@

\begin{mainth}
\label{thm:strip}
If $s_1$ and $s_2$ are the subgoals of the goal $φ$, that is,
$\fstrip~φ~n~≡~s_1\wedge~s_2$.
If $Γ ⊢ s_1$ and $Γ ⊢ s_2$ then $Γ ⊢ φ$.
\end{mainth}

>>>>

\begin{mainlemma}
  \label{lem:inv-uh-lem}
Let $\text{n} : \Nat$ be the complexity measure of the \rm{\fuh₀} function
in~(\ref{eq:strip-unfixed}).
If $Γ ⊢ \fuh₁~φ~n$ then $Γ ⊢ φ$.
\end{mainlemma}

\begin{proof}
Use induction on the cases defined by the outcome of the
$\fuh$ function in~(\ref{eq:uh-structured}).
\begin{itemize}
  \item If $n = 0$, by definition we conclude $Γ ⊢ φ$.
  \item If $n = 1$, we apply the following theorem proved in~\cite{AgdaProp},
    \begin{equation*}
    \texttt{∧⇒\rm{-to-}⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
    \end{equation*}
  \item For $n > 1$, we use induction on the structure of the second
        argument.
\vskip 2mm
\begin{itemize}

@@

Finally, using the theorem \texttt{⇒∧⇒\rm{-to-}⇒∧} from~\cite{AgdaProp},
\begin{equation*}
  \texttt{⇒∧⇒\rm{-to-}⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃),
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}_1$}
  \AxiomC{$\mathcal{D}_2$}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
  \RightLabel{\tt ⇒∧⇒\rm{-to-}⇒∧}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)}
  \end{bprooftree}
\end{equation*}
\item Other cases are proved in a similar way.
\end{itemize}
\end{itemize}
\end{proof}

@@

%  1137  git add .
%  1138  git commit -am "[ paper ] strip proof."
%  1139  git pull origin paper
%  1140  git push origin paper
\begin{equation}
\label{eq:strip-fixed}
\begin{aligned}
&\hspace{.495mm}\fstrip : \Prop → \Nat → \Prop\\
&\begin{array}{llll}
\fstrip &(φ₁ ∧ φ₂)     &(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ∨ φ₂)     &(\suc~n) &= \fuh~(¬ φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ⇒ φ₂)     &(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(¬ (φ₁ ∧ φ₂)) &(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ∨ φ₂)) &(\suc~n) &= \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ⇒ φ₂)) &(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ ¬ φ₁)      &(\suc~n) &= \fuh~(\fstrip~φ₁~n)\\
\fstrip &(¬ ⊥)         &(\suc~n) &= ⊤\\
\fstrip &(¬ ⊤)         &(\suc~n) &= ⊥\\
\fstrip &φ             &\text{n} &= φ
\end{array}
\end{aligned}
\end{equation}

\begin{mainlemma}
\label{lem:lem-inv-strip}
Let $\text{n} : \Nat$ be the complexity measure of the strip function
in~(\ref{eq:strip-unfixed}).
If $Γ ⊢ \fstrip~φ~n$ then $Γ ⊢ φ$.
\end{mainlemma}

\begin{proof}
The proof is by induction on the structure of the
formula~$φ$ by following the cases in~(\ref{eq:strip-fixed}).
We present a straightforward case with double negation, a case with
conjunction connective, and last, the case with a negated disjunction.
We refer the reader to~\cite{AgdaMetis} for the complete proof in \Agda.

$\bullet$~Case $φ ≡ ¬ ¬ φ₁$.
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \fstrip~(¬¬φ₁)~(\suc~n)$}
  \RightLabel{by~(\ref{eq:uh-fixed})}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n)$}
  \RightLabel{Lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ \fstrip~φ₁~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

$\bullet$~Case $φ ≡ φ₁ ∧ φ₂$. We get a proof for each conjunct and using the
introduction rule for conjunction connective, the expected result follows.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fstrip~(φ₁ ∧ φ₂)~(\suc~n)$}
  \RightLabel{by~(\ref{eq:strip-fixed})}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n)$}
  \RightLabel{Lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ \fstrip~φ₁~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ φ₁$}
    \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \fstrip~(φ₁ ∧ φ₂)~(\suc~n)$}
  \RightLabel{by~(\ref{eq:strip-fixed})}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{Lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip~φ₂~n$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \fstrip~φ₂~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ φ₂$}
  \end{bprooftree}
\end{equation*}

$\bullet$~Case $φ ≡ ¬ (φ₁ ∨ φ₂)$.
We show the theorems $Γ ⊢ ¬ φ₁$ and $Γ ⊢ ¬ φ₂$.
With the conjunction of $¬ φ₁$ and $¬ φ₂$ by applying De Morgan Law, the
result follows.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \fstrip~(¬ (φ₁ ∨ φ₂))~(\suc~n)$}
\RightLabel{by~(\ref{eq:strip-fixed})}
\UnaryInfC{$Γ ⊢ \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ \fuh~(\fstrip~¬ φ₁~n)$}
\RightLabel{Lemma~\ref{lem:inv-uh-lem}}
\UnaryInfC{$Γ ⊢ \fstrip~¬ φ₁~n$}
\RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ ¬φ₁$}
\end{bprooftree}
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{$Γ ⊢ ¬ φ₁$}
  \AxiomC{$Γ ⊢ \fstrip~(¬ (φ₁ ∨ φ₂))~(\suc~n)$}
  \RightLabel{by~(\ref{eq:strip-fixed})}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)$}
  \RightLabel{∧-proj₂}
  \UnaryInfC{$Γ ⊢ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)$}
  \RightLabel{Lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \fstrip~¬ φ₂~n$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ ⊢ \fstrip~¬ φ₂~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ ¬ φ₂$}
  \end{bprooftree}
\end{equation*}
\end{proof} % the proof continue...

MISSING TEXT

@@

\begin{mainth}
\label{thm:strip}
If $s_1$ and $s_2$ are the subgoals of the goal $φ$, that is,
$\fstrip~φ~n~≡~s_1\wedge~s_2$, if $Γ ⊢ s_1$ and $Γ ⊢ s_2$ then $Γ ⊢ φ$.
\end{mainth}
